Pipe-based Programming

### 2018-11-19 doc/paper.tex
update paper

@@

\title{Concurrency for Pipe-based Programming}
\author{Jeanine Adkisson}
\affiliation{\institution{Tokyo Institute of Technology}}
\email{jneen@prg.is.titech.ac.jp}

@@

\begin{abstract}
The UNIX shell programming model has been shown to be a successful model for composing programs - spawning unrelated programs in parallel to communicate over operating-system pipes. Beyond the most simple tasks, however, \texttt{bash} and similar tools break down, due to the byte-based nature of the system: variables can only hold byte arrays, and pipes can only manage raw byte streams, with no support for transactional writes.

We explore a pipe-based programming model that attempts to synthesize the good parts of UNIX shells, departing at key points, along with a supporting concurrency model and a semantic of interruption.

@@

>>>>

% from matias
\usepackage{balance}

@@

\title{Pipe-based Programming}
\author{Jeanine Adkisson}
\affiliation{\institution{Tokyo Institute of Technology}}
\email{jneen@prg.is.titech.ac.jp}

@@

\begin{abstract}
Shell languages such as bash are designed to integrate with an OS, which mainly involves managing processes with implicit input and output streams. They also attempt to do this in a compact way that could be reasonably typed on a command-line interface.

However, existing shell languages are not sufficient to serve as general-purpose languages - values are not observable except in raw streams of bytes, and they lack modern language features such as lexical scope and higher-order functions.

By way of a new programming language, \emph{Magritte}, we propose a general-purpose programming language with semantics similar to bash. In this paper, we discuss the early design of such a system, in which the primary unit of composition, like bash, is processes with asynchronous inputs and outputs, which can be read from or written to at any time, and which can be chained together via a pipe operator. We also explore concurrency semantics for such a language.

We explore the implementation and design challenges that come with such a model, and also show that this model is able to easily create and compose concurrent algorithms.
The UNIX shell programming model has been shown to be a successful model for composing programs - spawning unrelated programs in parallel to communicate over operating-system pipes.

% Beyond the most simple tasks, however, \texttt{bash} and similar tools break down, due to the byte-based nature of the system: variables can only hold byte arrays, and pipes can only manage raw byte streams, with no support for transactional writes.
% %
% We explore a pipe-based programming model that attempts to synthesize the good parts of UNIX shells, departing at key points, along with a supporting concurrency model and a semantic of interruption.
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Functional and object-oriented programming suffer from a disconnect between the concept of \emph{functions} and \emph{programs}.  Whereas functions and methods receive values as arguments and return single values as a result, programs can receive data not only as arguments but also asynchronously through an input, and write results asynchronously to an output.
%
% In functional and object-oriented languages, this can result in not only a significant amount of abstraction overhead for dealing with external functions, but the immediacy of function arguments and return values can make composing concurrent programs difficult.
%
% By way of a new programming language, \emph{Magritte}, we propose a programming model in which the primary unit of composition is not immediate function arguments and return values, but instead processes with asynchronous inputs and outputs, which can be read from or written to at any time, and which can be chained together via a pipe operator, much like programs in a UNIX environment. We also explore a concurrency semantic for such a language.
%
% We explore many implementation and programming challenges that come with such a model, and also show that this model is able to easily create and compose concurrent algorithms that would be difficult to write or compose in a functional or object-oriented language.

@@

%# fn pipe(lhs, rhs, inp, out) {
%#   c = channel()
%#   go fn () { lhs(inp, c); };
%#   rhs(c, out)
%# }
%#
%# collect = collector()
%# pipe_collect(curry(range, 10), curry(fan, 2,
%  fn(el, inp, out) { write(el + 1, out); }
%), inp);
%#
